XXX

### 2012-03-24 paper/engNERwiki.tex
Stuff.

\section{Introduction}

Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as Information Retrieval, Information Extraction or Machine Translation. The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995 \cite{Grishman:96}. In MUC shared tasks the NER consists of three subtasks: entity names, temporal expressions and number expressions. However there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations. The fourth type, called "miscellaneous'', is used in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types. Since then, MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.

@@

One approach to reduce the manual intervention required in language resource production, and thus ultimately production costs is automatically generating resources. An other approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data. Researchers apply the Web-as-corpus approach, often dealing with such datasets as the Google 5G 1T collection or DBPedia.

@@

\section{Wikipedia and NER}
% Wikify!
% stuff from Mining Meaning from Wikipedia
%NER and Wikipedia

Wikipedia is a goldmine of information; it is applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual Information Retrieval, and Question Answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.

@@

To our knowledge, such type of automatically built corpora are not freely available, with the only one exception of the Semantically Annotated Snapshot of the English Wikipedia \cite{Zaragoza:07}. It provides a wide range of linguistic information: POS tags, dependency labels, WordNet super senses and NE annotation according to WSJ and CoNLL tagsets. However, they followed a different way: the NEs were tagged by open source softwares. Our corpora will be available under the same license as Wikipedia's one (Creative Commons Attribution-Sharealike 3.0 Unported License).

Thanks for the huge amount of Wikipedia articles, the possibility of building large enough corpora is given for less resourced languages, as well. Since our method is mainly language-independent, it can be easily applied for languages which are structurally differing from English. Hungarian is a highly agglutinative language, with free word order and other particularities detailed later in section \ref{sec:hun}. There are human annotated CoNLL-style datasets, but this presented here is the first large enough automatically NE-annotated corpus for Hungarian.

@@

% Introduction: Wikipedia + DBpedia

Wikipedia proves to be a rich source of information about its entities; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure. Since our goal is to create a NE-annotated corpus containing large amount of texts, our interest is limited to the article text. Templates are removed from the articles; however, the names of the templates have been retained for XXX, but all data (including infobox)

Beside regular articles, Wikipedia also contains redirect and disambiguation pages, which require special treatment. Redirect pages are empty, and are safe to ignore altogether. Disambiguation pages contain text, usually in the form of an enumeration of page links with a short description. Full sentences are scarce, and the structure of these pages makes it hard to extract valid training sentences from them. Therefore, disambiguation pages are dropped from the corpus, as well.

The English corpus is based on the Wikipedia snapshot as of XXX. The XML files were parsed with the mwlib (CITE) parser. The raw text was tokenized with a modified version of the XXX sentence tokenizer in NLTK (CITE). The words were lemmatized with XXX and POS tagged with HunPos \cite{Halacsy:07}.

@@

@@

\subsection{The DBpedia -- ConLL mapping}
\label{dbpedia}
% Introduction + entity -> class -> category.

@@

@@

>>>>

\section{Introduction}
\label{sec:intro}

Named Entity Recognition (NER), the task of identifying Named Entities (NEs) in unstructured texts and classifying them into pre-selected classes, is one of the most important subtasks in many NLP tasks, such as Information Retrieval, Information Extraction or Machine Translation. The NER task was introduced with the 6th Message Understanding Conference (MUC) in 1995 \cite{Grishman:96}. In MUC shared tasks the NER consists of three subtasks: entity names, temporal expressions and number expressions. Although there is a general agreement in the NER community about the inclusion of temporal expressions and some numerical expressions such as amounts of money and other types of units, the most studied types are names of persons, locations and organizations. The fourth type, called "miscellaneous'', was introduced in the CoNLL NER tasks in 2002 \cite{Tjong:02} and 2003 \cite{Tjong:03}, and includes proper names falling outside the three classic types. Since then, MUC and CoNLL datasets and annotation schemes have been the major standards applied in the field of NER.

@@

One approach to reduce the manual intervention required in language resource production, and thus ultimately production costs is automatically generating resources. An other approach to achieve this aim is using collaborative annotation and/or collaboratively constructed resources, such as Wikipedia, Wiktionary or Linked Open Data. Researchers apply the Web-as-corpus approach, often dealing with such datasets as Google's Web 1T 5-gram collection or DBPedia.

@@

\section{Wikipedia and NER}
\label{sec:related}
% Wikify!
% stuff from Mining Meaning from Wikipedia
%NER and Wikipedia

Wikipedia is a free, multilingual Internet encyclopedia, written collaboratively by volunteers. (CITE http://en.wikipedia.org/wiki/Wikipedia). It is a goldmine of information; at the time of writing, Wikipedia contains about 21 million interlinked articles. Of these, 3,903,467 are English and 212,120 are Hungarian. XXX

Wikipedia has been applied for several NLP tasks, e.g.~word sense disambiguation, ontology and thesaurus building, multilingual Information Retrieval, and Question Answering, as Medelyan et al.~\shortcite{Medelyan:09} point out in their fully detailed survey about mining meaning from Wikipedia.

@@

To our knowledge, such type of automatically built corpora are not freely available, with the one exception of the Semantically Annotated Snapshot of the English Wikipedia \cite{Zaragoza:07}. It provides a wide range of linguistic information: POS tags, dependency labels, WordNet super senses and NE annotation according to WSJ and CoNLL tagsets. However, they followed a different way: the NEs were tagged by open source softwares. Our corpora is available under the same license as Wikipedia's one (Creative Commons Attribution-Sharealike 3.0 Unported License).

Thanks for the huge amount of Wikipedia articles, the possibility of building large enough (XXX ez mit jelent?) corpora is given for less resourced languages, as well. Since our method is mainly XXX language-independent, it can be easily applied for languages which are structurally differing from English. Hungarian is a highly agglutinative language, with free word order and other particularities XXX detailed later in section \ref{sec:hun}. There are human annotated CoNLL-style datasets, but the one presented here is the first large enough automatically NE-annotated corpus for Hungarian.

@@

Our goal is to create a large NE-annotated corpus, automatically generated from Wikipedia articles. We follow a similar path XXX to \cite{Nothman:08}, who broke up XXX the process into four steps:
\begin{enumerate}
\item Classify all articles into entity classes
\item Split Wikipedia articles into sentences
\item Label \texttt{NE}s according to link targets
\item Select sentences for inclusion in a corpus
\end{enumerate}

In this section, we describe how these steps were implemented. While the approach in essence is language-agnostic, the exact details XXX have to be accomodated to the target language. This section explains the general idea and its execution for English; the next section describes how the idea can be adapted to Hungarian.

\subsection{Parsing Wikipedia}

Wikipedia is very rich source of information; aside from the article text, a huge amount of data is encompassed in infoboxes, templates and the category structure. Our task requires the use of only the links between the articles and the article text. In addition to in-article links, our method takes advantage of the redirect and interlanguage links, available as SQL dumps.

Beside regular articles, Wikipedia also contains redirect and disambiguation pages, which require special treatment. Redirect pages are empty, and are safe to ignore altogether. Disambiguation pages contain text, usually in the form of an enumeration of page links with a short description. Full sentences are scarce, and the structure of these pages makes it hard to extract valid training sentences from them. Therefore, disambiguation pages are dropped from the corpus as well.

The English corpus is based on the Wikipedia snapshot as of January 15, 2011. The XML files were parsed with the \texttt{mwlib} (CITE http://code.pediapress.com) parser. The raw text was tokenized with a modified version of the Punkt sentence and word tokenizers \cite{Kiss:06} in NLTK \cite{Loper:02}. The words were lemmatized with the Wordnet lemmatizer, also in NLTK, and POS tagged with HunPos \cite{Halacsy:07}.

@@

\subsection{Articles as Entities}
\label{dbpedia}

Most authors

@@

@@

\bibitem[\protect\citename{Loper \bgroup et al.\egroup}2002]{Loper:02}
Edward Loper, Steven Bird.
\newblock 2002.
\newblock NLTK: the Natural Language Toolkit. 
\newblock In: {\em Proceedings of the ACL-02 Workshop on Effective tools and methodologies for teaching natural language processing and computational linguistics}, Volume 1, pages 63--70.

\bibitem[\protect\citename{Loper \bgroup et al.\egroup}2002]{Kiss:06}
Tibor Kiss, Jan Strunk.
\newblock 2006.
\newblock Unsupervised Multilingual Sentence Boundary Detection. 
\newblock In: {\em Computational Linguistics}, 32 (4): pages 485--525.

@@

% Jobban ki kell hangsúlyozni az introductionben, és a related worksben, hogy szabadon elérhető training korpuszt építettünk
% És azt, hogy Hungarian middle, English nagy.
